---
title: "PA1_Template"
author: "Natalia Flechas"
date: "19/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **View Data**

```{r}
activity<-read.csv("activity.csv",header = TRUE,sep = ",")
names(activity)
str(activity)
activity$date<-as.Date(activity$date)
```

## Charge Packages

```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
theme_set(theme_light())
```

## Transform data

```{r}
weekday<-weekdays(activity$date)
activity<-cbind(activity,weekday)
head(activity)
activity<-as_tibble(activity)
str(activity)
```

## To view summary data 

```{r}
summary(activity)
```

## Histogram of steps per day 


```{r}
histogram_steps_per_day1<-activity%>%select(steps,date)%>%group_by(date)%>%summarise(sumSteps=sum(steps))
a<-ggplot(histogram_steps_per_day1,aes(x=sumSteps))
b<-a+geom_histogram(bins=30,color="black",fill="yellow")+labs(title = "Histogram of steps per day",x="Steps per day",y="Frequency")
b
```

## Mean and Median of data 

```{r}
histogram_steps_per_day1<-na.omit(histogram_steps_per_day1)
mean(histogram_steps_per_day1$sumSteps)
median(histogram_steps_per_day1$sumSteps)
```

## **What is the average daily activity pattern?**

```{r}
daily_activity_pattern<-aggregate(activity$steps,by=list(activity$interval),FUN=mean,na.rm=TRUE)
daily_activity_pattern<-as_tibble(daily_activity_pattern)
ts_dailypattern<-ggplot(data = daily_activity_pattern,aes(x=Group.1,y=x))+geom_line(color="#00AFBB",size=1)+labs(title="Time series of steps by interval",x="Interval",y="mean of steps")
ts_dailypattern
```

## Maximum of number of steps by interval 

```{r}
max(daily_activity_pattern$x)
```

## **Input missing values**

## Calculate and report of total number of missing values in dataset 
```{r}
sum(is.na(activity$steps))
````

## The strategy for imput missing values is not sophisticated. For example we can use the mean/median for data day. 

```{r}
activity_impute_mean<-data.frame(sapply(activity,function(x)ifelse(is.na(x),mean(x,na.rm = TRUE),x)))
```

## See the histogram of imputed data 

```{r}
activity_impute_mean$steps<-as.numeric(activity_impute_mean$steps)
activity_impute_mean$interval<-as.integer(activity_impute_mean$interval)
Hactivity_impute_mean<-activity_impute_mean%>%select(date,steps)%>%group_by(date)%>%summarise(sumSteps=sum(steps))
head(Hactivity_impute_mean)
b<-ggplot(Hactivity_impute_mean,aes(x=sumSteps))
c<-b+geom_histogram(bins = 30,color="black",fill="blue")+labs(title = "Histogram of steps in imputed data",x="Steps",y="Frequency")
c
```

## Mean and Median of imputed data

```{r}
mean(Hactivity_impute_mean$sumSteps)
median(Hactivity_impute_mean$sumSteps)
```

## Imputed data has not difference with initial data. 

## **Are there differences in activity patterns between weekdays and weekends?**

### Weekday variable.
```{r}
 innerweek<-activity_impute_mean%>%select(steps,interval,weekday)%>%filter(weekday%in%c("lunes","martes","miércoles","jueves","viernes"))
head(innerweek)
```

### Plot weekday time series.
```{r}
innerweek<-activity_impute_mean%>%select(steps,interval,weekday)%>%filter(weekday%in%c("lunes","martes","miércoles","jueves","viernes"))
innerweek$steps<-as.numeric(innerweek$steps)
innerweek$interval<-as.integer(innerweek$interval)
str(innerweek)
ts_innerweek<-aggregate(innerweek$steps,by=list(innerweek$interval),FUN=mean)
names(ts_innerweek)<-c("interval","steps")
```
```{r}
q<-ggplot(data = ts_innerweek,aes(x=interval,y=steps))+geom_line(color="#DF0174",size=1)+labs(title="Time series of steps per interval in weekdays",x="Interval of minutes",y="Steps")
q
```

### Summary of weekday data 
```{r}
summary(ts_innerweek$steps)
```

### Weekend variable.
```{r}
weekends<-activity_impute_mean%>%select(steps,interval,weekday)%>%filter(weekday%in%c("sábado","domingo"))
```

### Plot weekend time series. 
```{r}
weekends$steps<-as.numeric(weekends$steps)
weekends$interval<-as.integer(weekends$interval)
ts_weekends<-aggregate(weekends$steps,by=list(weekends$interval),FUN=mean)
names(ts_weekends)<-c("interval","steps")
r<-ggplot(ts_weekends,aes(x=interval,y=steps))+geom_line(color="#01DF3A",size=1)+labs(title = "Time series of steps per interval in weekends",x="Interval of minutes",y="Steps")
r
```

### Summary of weekend data.
```{r}
 summary(ts_weekends$steps)
```








